---
title: "Cleaning and Wrangling Data"
author: "Yulissa T. Perez Rojas"
date: "2023-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r load packages, warming=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

# Introduction

About the data: This data was downloaded from [this link](https://knb.ecoinformatics.org/view/df35b.304.2)

## Read data

```{r read data, warning=FALSE, message=FALSE}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

```

## Explore data

```{r explore }
colnames(catch_original)

head(catch_original)

summary(catch_original)

#Prints unique values in a year and sort them.  
sort(unique(catch_original$Year))

#how many regions are in the data? 
unique(catch_original$Region)

```

# Clean and Reshape data

## Selecting and Removing columns with `select()`

```{r select}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

# Adding a minus is to NOT select specific columns 
catch_data <- catch_original %>% 
  select(-All, - notesRegCode)

```

## Check column types

```{r checking}

glimpse(catch_data)

```

## Changing column content using `mutate()`

```{r changing data type}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

```

## Checking NA

```{r check NA}
# Where is the NA in the Chinook column?
i <- which(is.na(catch_clean$Chinook))

# Inspect the actual line 
catch_data[i,]
```

# Fixing Chinook column

```{r fix column}
# Override column value (mutate), if its "I" (if_else) then (,) replace with 1. Then make "1" to a numeric value (as.numeric)
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", 
                           "1",
                           Chinook),
         Chinook = as.numeric(Chinook))

# Check if the code worked
catch_clean[i,]

```

## Reshape data

```{r reshape}
# Reshape columns by using pivot_long but not certain columns (-) then (,) create two new columns that contains species (names_to) and another one that contains numbers(values_to)
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year),
               names_to = "Species",
               values_to = "Catch")

head(catch_long)

```

For long to wide use `pivot_wider()`

```{r reshape wide}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Species,
              values_from = Catch)

head(catch_wide)

```

## Renaming column

```{r rename column}
catch_long <- catch_long %>% 
  rename(catch_thousand = Catch)

colnames(catch_long)

```

## Add a column using `mutate()`

```{r adding column}
catch_long <- catch_long %>% 
  mutate(Catch = catch_thousand*1000) %>% 
  select(-catch_thousand)

head(catch_long)

```

## Split-Apply-Combine workflow

We are going to use `group_by()` and `summarize()`

```{r split-apply-combine}
# Find the mean of the catch based on group. Group first then find mean.
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(catch_mean = mean(Catch))

head(mean_region)

# Now group two tags and find the mean
mean_reg_year <- catch_long %>% 
  group_by(Region, Year) %>% 
  summarise(catch_mean = mean(Catch))

head(mean_reg_year)

#Count how many rows are in that group
n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(obs_number = n())

head(n_region)

```

## Filtering by rows with `filter()`
```{r filtering}

sse_catch <- catch_long %>% 
  filter(Region == "SSE")

head(sse_catch)

```


## All together
```{r}
catch_original <- read_csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"))


mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), 
               names_to = "species", 
               values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch))

head(mean_region)

```

