---
title: "Intro to Data Visualization"
author: "Yulissa T. Perez Rojas"
date: "2023-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r load packages, warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(forcats)
library(ggplot2)
library(leaflet)
library(DT)
library(scales)

```

## Read data
```{r  read data, warning=FALSE, message=FALSE}
escape <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e")
```

## About data
This data was collected (here)[https://knb.ecoinformatics.org/view/doi%3A10.5063%2FP26WJJ]

## Explore data
```{r explore}

colnames(escape)

head(escape)

glimpse(escape)

range(escape$sampleDate)

unique(escape$Species)

```


## Getting data ready to plot
```{r wrangling}

annual_esc <- escape %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

unique(annual_esc$Species)


```

## Plotting with `ggplot()`

## Essential components
```{r essen}

##option 1 
ggplot(data = annual_esc,
       aes(x = Species, y = escapement)) +
    geom_col()

##option 2
ggplot(data = annual_esc) +
    geom_col(aes(x = Species, y = escapement))

##option 3
ggplot() +
    geom_col(data = annual_esc,
             aes(x = Species, y = escapement))

```


**Different geoms**
```{r plot}

annual_esc %>% 
  filter(Year == 1974,
         Species %in% c("Chum", "Pink")) %>% 
  ggplot(aes(x = Species,
             y = escapement)) +
  geom_boxplot()

```


**Violin Plot**
```{r violin}

annual_esc %>%
    filter(Year == 1974,
           Species %in% c("Chum", "Pink")) %>%
    ggplot(aes(x = Species, y = escapement)) +
    geom_violin()

```

**Line and Point**
```{r line}
annual_esc %>%
    filter(Species  == "Sockeye",
           SASAP.Region == "Bristol Bay") %>%
    ggplot(aes(x = Year, y = escapement)) +
    geom_line() +
    geom_point()

```


### Customizing plot
Fill = "blue" outside `aes()` to actually color the columns. If its inside the eas function, it will not work since it think it's a variable
```{r custom}

ggplot(data = annual_esc,
       aes(x = Species,
           y =  escapement)) +
  
  geom_col(fill = "blue")

```

Adding color by region
```{r region color}

ggplot(annual_esc,
       aes(x = Species, y = escapement,
           fill = SASAP.Region)) +
    geom_col()

```


### Creating multiple plots using `face_wrap`

Subset data
```{r subset}
annual_esc_2000 <- annual_esc %>% 
  filter(Year %in% c(2000:2016))

#quick checking
unique(annual_esc_2000$Year)

```

Plot with `face_wrap`
```{r facelift}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y")

```


### Setting ggplot themes
`theme*()`
```{r theme function}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y") +
  theme_bw() +
   ylab("Escapement") +
  ggtitle("Annual Salmon Escapement by Region")
```

Removing legend title and moving to bottom of plot
```{r edit plot vis}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y") +
  theme_bw() +
   ylab("Escapement") +
  ggtitle("Annual Salmon Escapement by Region") +
  theme(legend.position = "bottom") +
  theme(legend.title = element_blank())

```

### Saving theme into an object
```{r saving theme}

my_theme <- theme_bw() +
  theme(legend.position = "bottom") +
  theme(legend.title = element_blank())   

```

Creating plot with saved theme
```{r plot w/ theme}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y") +
  ylab("Escapement") +
  ggtitle("Annual Salmon Escapement by Region") +
  my_theme 
  
```

Angle years to 45 degree and break x-axis by 2
```{r angle break}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y") +
  ylab("Escapement") +
  scale_x_continuous(breaks = seq(2000,2016, 2)) + #breaking
  ggtitle("Annual Salmon Escapement by Region") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,     #rotation
                                     vjust = 0.5))
```

Changing scientific notation to commas in y-axis

```{r scale}

ggplot(annual_esc_2000,
       aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region,
             scales = "free_y") +
  ylab("Escapement") +
  scale_x_continuous(breaks = seq(2000,2016, 2)) + #breaking
  scale_y_continuous(labels = comma)               #no longer scientific fig.
  ggtitle("Annual Salmon Escapement by Region") +
  theme(axis.text.x = element_text(angle = 45,     #rotation
                                     vjust = 0.5)) +
  my_theme
  
```

Creating an object of the plot 
```{r plot object}

annual_region_plot <- ggplot(annual_esc_2000,
                             aes(x = Year,
                                 y = escapement,
                                 color = Species)) +
    geom_line() +
    geom_point() +
    scale_x_continuous(breaks = seq(2000, 2016, 2)) +
    scale_y_continuous(labels = comma) +
    facet_wrap( ~ SASAP.Region,
                scales = "free_y") +
    ylab("Escapement") +
    xlab("\nYear") +
    ggtitle("Annual Salmon Escapement by Region") +
    my_theme +
    theme(axis.text.x = element_text(angle = 45,
                                     vjust = 0.5))
```

### Saving plot
```{r save plot}

ggsave("figs/annual_esc_region.png", annual_region_plot, , width = 12, height = 8, units = "in")

```


## Maps with `leaflet()`

Subset location information
```{r subet location info}

location <- escape %>% 
  distinct(Location, Latitude, Longitude) %>%  #These variables are in data
  drop_na()

```

Create map with `leaflet()`
```{r map location}

leaflet(location) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude,
             lat = ~ Latitude,
             popup = ~ location)
```

Map with different base map
```{r map with diff base}

leaflet(location) %>%
    addWMSTiles(
        "https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?request=getmap&service=wms&BBOX=-90,-180,90,360&crs=EPSG:4326&format=image/jpeg&layers=gebco_latest&width=1200&height=600&version=1.3.0",
        layers = 'GEBCO_LATEST',
        attribution = "Imagery reproduced from the GEBCO_2022 Grid, WMS 1.3.0 GetMap, www.gebco.net"
    ) %>%
    addCircleMarkers(
        lng = ~ Longitude,
        lat = ~ Latitude,
        popup = ~ Location,
        radius = 5,
        # set fill properties
        fillColor = "salmon",
        fillOpacity = 1,
        # set stroke properties
        stroke = T,
        weight = 0.5,
        color = "white",
        opacity = 1
    )

```


