---
title: "Introduction to RMarkdown"
author: "Yulissa T. Perez Rojas"
date: "2023-04-03"
output: 
  html_document: 
    toc: true
    toc_float: true
bibliography: references.bib
reference-section-title: Reference
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Markdown

## This is a header

-   This is **bold**
-   This is *italic*

## Syntax for my reseach

-   To write subscript use the \~ so CO~2~
-   Add links by using the brackets [Link]
    -   Link to AmeriFlux [<https://ameriflux.lbl.gov/>]
-   Adding citations, use the Visual tab instead of the Source tab then click insert.
    -   Click citation and add DOI [@oke2020]

## Adding Tables

-   To add tables, work on the Visual tab because it creates the syntax structure

    | Col1 | Col2 | Col3 |
    |------|------|------|
    |      |      |      |
    |      |      |      |
    |      |      |      |

: This is an example

## R Code chunks

-   Window shortcut is Control + Alt + i
-   Press the forward error to run code or use Control + Enter

```{r example}
heights_ft <- c(5.2, 6.0, 5.7)

important_coef <- 3.14


```

# Exercise: New Chunck

```{r}
cat_years <- c(2, 6, 12, 1)

mean(cat_years)
```
