---
title: "Intro to sf package for spatial data and making maps"
author: "Yulissa T. Perez Rojas"
date: "2023-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r load, warning=FALSE, message=FALSE}
library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap) # base maps to static maps
library(dplyr)

```

## Working Data

This data was downloaded from this [link](https://dev.nceas.ucsb.edu/view/urn:uuid:6f07cb25-a4a1-48e8-95cb-74f532f3ce2d)

```{r upload data}

ak_regions <- read_sf("data/ak_regions_simp.shp")

```

## Explore AK Regions data

```{r explore}

plot(ak_regions)

head(ak_regions)

glimpse(ak_regions)

```

## Information about CRS

```{r crs info}

st_crs(ak_regions)

```

### Transform CRS

Need to unprojected to projected geometry to be able to plot in a flat surface

```{r transform}

ak_regions_3338 <- ak_regions %>% 
  st_transform(crs = 3338)

# Check transformation
plot(ak_regions_3338)

```

## sf & tidyverse

```{r sf & tidyverse}

colnames(ak_regions_3338)

ak_regions_3338 %>% 
  select(region)

```

From all the regions only pick one using `filter()`

```{r filter}

ak_regions_3338 %>% 
  filter(region == "Yukon")

```

## Spatial Join

How many people live in each of these Alaska regions? (5 Steps)

```{r pop upload}

pop <- read_csv("data/alaska_population.csv")

class(pop)

```

Step 1: The population data is in a table (data.frame) and for mapping it needs to be spatial object. To do so, use `st_as_sf ()`

```{r pop data change}

pop_4326 <- st_as_sf(pop,
                     coords = c("lng", "lat"),
                      crs = 4326,
                     remove = FALSE)
class(pop_4326)
```

Step 2: Now that population is in a spatial object (unprojected). Next, step is to transform it to projected in order for all spatial objects to be the same type (projected).

```{r pop transform}
pop_3338 <- st_transform(pop_4326, crs = 3338)

st_crs(pop_3338)
```

Step 3: We need to join the population data with region geometry by using `st_join()` In addition, we are using the `st_within()` function to get the total populations the regions

```{r join pop and region}

pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

```

Step 4: To use `group_by()` function, need to change spatial object to data.frame.

```{r join pop and region pt.2 }

pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population)) 

```

Step 5: Use a regular `left_join()` to get the information back to the Alaska region **shapefile**.

```{r pop ~ region}

pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")

plot(pop_region_3338["total_pop"])

```

Total population by management data instead of region.

```{r pop ~ managment area}

pop_mgmt_338 <- pop_region_3338 %>%
    group_by(mgmt_area) %>%
    summarize(total_pop = sum(total_pop), do_union = FALSE) #preserved regions

plot(pop_mgmt_338["total_pop"])

```

## Save spatial object to a new file

Use `write_sf()`

```{r save to new file}

write_sf(pop_region_3338, "data/ak_regions_population.shp")
```

## Visualize with ggplot

```{r plot pop ~ region}

ggplot(pop_region_3338) +
  geom_sf(aes(fill = log10(total_pop))) +
  labs(fill = "Total Population") +
  scale_fill_viridis_c(labels = comma)
  

```

Adding rivers to data

```{r add rivers}

rivers <- read_sf("data/ak_rivers_simp.shp")

st_crs(rivers)

# Added 3338 after seeing the rivers shapefile is the same as alaska type "Alberts"
rivers_3338 <- rivers

```

Plot with multiple data objects

```{r plot layering}

ggplot()+
  geom_sf(pop_region_3338, mapping=aes(fill = total_pop))+
  geom_sf(pop_3338, mapping=aes())+
  geom_sf(rivers_3338, mapping=aes())

```

```{r fix plot layering}

ggplot()+
  geom_sf(pop_region_3338, mapping=aes(fill = total_pop))+
  geom_sf(pop_3338, mapping=aes(), color = "red", size = 0.7)+
  geom_sf(rivers_3338, mapping=aes(linewidth = StrOrder))+ #mess with river linewidth
  scale_linewidth(range = c(0.05, 0.5), guide = "none")+ #making river lines skinny
  scale_fill_viridis_c(labels = comma)+
  labs(fill = "Total Population",
       title = "Alaska Map Population")+
  theme(
    plot.title = element_text(hjust = 0.5))
  

```

# Leaflet Interactive Map

Using the `leafletCRS()` function allows you to creating a custom CRS for your data.

```{r leaflet}

epsg3338 <- leaflet::leafletCRS(
    crsClass = "L.Proj.CRS",
    code = "EPSG:3338",
    proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
    resolutions = 2 ^ (16:7)
)

```

Transform projected (Alberts) to unprojected because for interactive map you need a 3D map for interactive map.

```{r 4326 transform}

pop_region_4326 <- pop_region_3338 %>% 
  st_transform(crs = 4326)

```

Creating interactive map

```{r interactive map}

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>% 
     addPolygons(data = pop_region_4326,
                fillColor = "lightgreen",
                weight = 1)
  
m
```
