---
title: "RMarkdown Analysis"
author: "Yulissa T. Perez Rojas"
date: "2023-04-03"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction
North Pole Environmental Observatory Bottle Chemistry data (https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

This data set includes diverse measurements of seawater chemistry along with supporting conductivity, temperature, and depth (CTD) data from the Arctic Ocean near the North Pole. Measurements were taken from sea ice platforms each April or May from 2000-2014. 

This data was downloaded on April 3, 2023.

### Read Data

#### Set Up
```{r set up}
library(readr)
```

```{r read data, warning=FALSE, message=FALSE}
bg_chem <- read_csv("data/BGchem2008data.csv")
```

#### Data exploration
```{r data exploration for viewer, eval=FALSE, echo=FALSE}
## column names 
colnames(bg_chem)

## structure of data
str(bg_chem)

## summary data
summary(bg_chem)

## Open data in a tab
View(bg_chem)

## Unique values in a column
unique(bg_chem$Date)
```

## Analysis

### Calculate summary statistics 
```{r calculate mean}
nitrate_mean <- mean(bg_chem$NO3)
nitrite_mean <- mean(bg_chem$NO2)
amn_mean <- mean(bg_chem$NH4)
phos_mean <- mean(bg_chem$P)
```

### Calculate mead Redfild ratio
Redfield ratio 16:1
```{r ratio calc}
ratio <-(amn_mean+ nitrate_mean + nitrite_mean)/phos_mean

```

### Plot Refiled ratio
```{r plot}
plot(bg_chem$P, bg_chem$NO3 + bg_chem$NO2 + bg_chem$NH4)

```

## Conclusions 

The Redfield ratio is `r round(ratio)`
